# (PART) Introduction {.unnumbered}

# Introduction {#intro}


## Ojectives
This book explains several things,

1. An overview of the project from a technical perspective.
2. The interaction between R packages developed to manage the data and do the calculations.
3. The different types of data in the PIP project and the interaction between them.
4. How the poverty calculator, the table maker, and the Statistics Online (SOL) platform are updated. 
5. Technical standalone procedure necessary for the execution of some parts of the project. 

## Technical requirements

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')

set.seed(1234)
options(digits = 3)

knitr::opts_chunk$set(
  comment = "#>",
  warning = FALSE
  #collapse = TRUE,
  #cache = TRUE,
  #out.width = "70%",
  #fig.align = 'center',
  #fig.width = 6,
  #fig.asp = 0.618,  # 1 / phi
  #fig.show = "hold"
)

options(dplyr.print_min = 6, 
        dplyr.print_max = 6)

ipkg <- utils::installed.packages()[,1]
if (!("emo"  %in% ipkg)){
  remotes::install_github("hadley/emo")
}


```

<!-- ::: {.rmdbox .rmdwarning} -->
<!-- DISCLAIMER: This book contains information relevant exclusively for the internal -->
<!-- Technical team of the PIP project. It has been made available to the public for -->
<!-- the sake of transparency, but its information has no use outside the internal -->
<!-- team. -->
<!-- ::: -->


You need to make sure the `bookdown` package is installed in your computer

```{r eval=FALSE}
install.packages("bookdown")

# or the development version
 devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is
defined by the first-level heading `#`.

Make sure to install the latest version of PIP R packages by typing the
following

```{r, eval=FALSE}
ipkg <- utils::installed.packages()[,1]

pip_install <- function(pkg, ipkg) {
  if (isFALSE(pkg %in% ipkg)) {
    gitcall <- paste0("PIP-Technical-Team/", pkg)
    remotes::install_github(gitcall, dependencies = TRUE)
    TRUE
  } else {
    FALSE
  }
}

pkgs <- c("pipload", "pipaux", "wbpip", "piptb", "pipdm", "pipapi")


purrr::walk(pkgs, pip_install, ipkg = ipkg)

```

