# Price FrameWork (pfw) data frame

The price framework data frame (pfw) is the most important source of technical
metadata of the PIP project, which makes pertinent the have of a separate
chapter for it. The general explanation of the structure of the pfw can be found
in Section \@ref(pfw-join). This chapter focuses more on the use of each of the
variables. Yet, it is worh repeating a few things already mentioned in Section
\@ref(pfw-join).

```{r}
pfw <- pipload::pip_load_aux("pfw")
```

## Variables

blah blah

### ID vars

Variables of identification are `wb_region_code`, `pcn_region_code`,
`country_code`, `ctryname`, `year`, `surveyid_year`, `survey_acronym`.

The main difference between `wb_region_code` and `pcn_region_code` is that the
former only include geographical regions internally used by the world bank,
whereas the latter has an additional category, "OHI," for Other High Income
countries.

```{r}
janitor::tabyl(pfw, wb_region_code, pcn_region_code )
```

The difference between `year` and `surveyid_year` is that... [Minh, could you
please explain here?]{style="color:red"}

### `fieldwork`

blah blah

### `altname`

blah blah

### `wbint_link`

blah blah

### `wbext_link`

blah blah

### `alt_link`

blah blah

### `surv_title`

blah blah

### `surv_producer`

blah blah

### `survey_coverage`

blah blah

### `welfare_type`

blah blah

### `use_imputed`

blah blah

### `use_microdata`

blah blah

### `use_bin`

blah blah

### `use_groupdata`

blah blah

### `reporting_year`

blah blah

### `survey_comparability`

blah blah

### `comp_note`

blah blah

### `preferable`

blah blah

### `show_portal`

blah blah

### `fieldwork_range`

blah blah

### `survey_year`

blah blah

### `newref`

blah blah

### `ref_year_des`

blah blah

### `wf_baseprice`

blah blah

### `wf_baseprice_note`

blah blah

### `wf_baseprice_des`

blah blah

### `wf_spatial_des`

blah blah

### `wf_spatial_var`

blah blah

### `cpi_replication`

blah blah

### `cpi_domain`

blah blah

### `cpi_domain_var`

blah blah

### `wf_currency_des`

blah blah

### `ppp_replication`

blah blah

### `ppp_domain`

blah blah

### `ppp_domain_var`

blah blah

### `wf_add_temp_des`

blah blah

### `wf_add_temp_var`

blah blah

### `wf_add_spatial_des`

blah blah

### `wf_add_spatial_var`

blah blah

### `tosplit`

blah blah

### `tosplit_var`

blah blah

### `inpovcal`

blah blah

### `oth_welfare1_type`

blah blah

### `oth_welfare1_var`

blah blah

### `gdp_domain`

blah blah

### `pce_domain`

blah blah

### `pop_domain`

blah blah

### `Note`

blah blah

### `pfw_id`

blah blah
