---
editor_options: 
  markdown: 
    wrap: 72
---

# (PART) PIP data {.unnumbered}

# Price FrameWork (pfw) data frame {#pfw-chapter}

The price framework data frame (pfw) is the most important source of
technical metadata of the PIP project, which makes pertinent the have of
a separate chapter for it. The general explanation of the structure of
the pfw can be found in Section \@ref(pfw-join). This chapter focuses
more on the use of each of the variables. Yet, it is worh repeating a
few things already mentioned in Section \@ref(pfw-join).

```{r}
pfw <- pipload::pip_load_aux("pfw")
```

## Variables

The table provides a short description of each variable of the Price
FrameWork data frame. We provide additional information for variable.

```{r, echo=FALSE}
pfwd <- readxl::read_xlsx(here::here("files/pfw_descriptions.xlsx"))
DT::datatable(pfwd)
```

## Additional explanation

### ID vars {.unnumbered}

Variables of identification are `wb_region_code`, `pcn_region_code`,
`country_code`, `ctryname`, `year`, `surveyid_year`, `survey_acronym`.

The main difference between `wb_region_code` and `pcn_region_code` is
that the former only include geographical regions internally used by the
world bank, whereas the latter has an additional category, "OHI," for
Other High Income countries.

```{r}
janitor::tabyl(pfw, wb_region_code, pcn_region_code)
```

For most of the data points, there should be no difference in between
`year` and `surveyid_year`. In some extreme cases (not available at the
moment), `year` could be different due to the reporting criteria from
the NSO. For our purpose, we use `rep_year` as the reporting year in our
system for integer ones, and `ref_year` for decimal reporting purpose

### `altname` {.unnumbered}

Alternative survey name for some surveys.

```{r}
head(pfw[altname != "", c("altname", "survey_acronym")])
```

### `survey_coverage` {.unnumbered}

This variable represent the househol survey coverage. This is different
from the disaggregation data level.

```{r}
janitor::tabyl(pfw, survey_coverage)
```

### `welfare_type` {.unnumbered}

This variable contains the welfare type of the main welfare aggregate
variable in a survey in case it has more than one. The welfare type of
alternative welfare aggregates is found in variable `oth_welfare1_type`.

```{r}
janitor::tabyl(pfw, welfare_type)
```

### `use_imputed` {.unnumbered}

Whether the welfare aggregate has been imputed. There are just few
countries with this kind of data.

```{r}
pfw[use_imputed == 1, unique(country_code)]
```

### `use_microdata` {.unnumbered}

Whether the welfare aggregate *vector* used in PIP is directly extracted
from microdata without any sort of aggregation. Mos of the countries
have this kind of data. Below you can see those that do not.

```{r}
pfw[use_microdata != 1, unique(country_code)]
```

### `use_bin` {.unnumbered}

Whether the welfare aggregate was aggregated to 400 bins from microdata
before being incorporated to the PIP repository. This is the case of
houshold surveys only available in the Luxembourg Data Center
([LIS](https://www.lisdatacenter.org/)). Countries with bin data is
considered micro data for technical purposes.

```{r}
pfw[use_bin == 1, unique(country_code)]
```

### `use_groupdata` {.unnumbered}

Whether welfare aggregate comes from grouped data. Information about
this type of data is available in [@dattComputationalToolsPoverty1998;
@krauseCorrigendumEllipticalLorenz2013;
@villasenorEllipticalLorenzCurves1989]. The following countries have
this kind of data.

```{r}
pfw[use_groupdata == 1, unique(country_code)]
```
