# (PART) PIP data {.unnumbered}

# Price FrameWork (pfw) data frame {#pfw-chapter}

The price framework data frame (pfw) is the most important source of technical
metadata of the PIP project, which makes pertinent the have of a separate
chapter for it. The general explanation of the structure of the pfw can be found
in Section \@ref(pfw-join). This chapter focuses more on the use of each of the
variables. Yet, it is worh repeating a few things already mentioned in Section
\@ref(pfw-join).

```{r}
pfw <- pipload::pip_load_aux("pfw")
```

## Variables

blah blah

### ID vars {-} 

Variables of identification are `wb_region_code`, `pcn_region_code`,
`country_code`, `ctryname`, `year`, `surveyid_year`, `survey_acronym`.

The main difference between `wb_region_code` and `pcn_region_code` is that the
former only include geographical regions internally used by the world bank,
whereas the latter has an additional category, "OHI," for Other High Income
countries.

```{r}
janitor::tabyl(pfw, wb_region_code, pcn_region_code)
```

The difference between `year` and `surveyid_year` is that... [Minh, could you
please explain here?]{style="color:red"}

### `altname` {-}

Alternative survey name for some surveys.

```{r}
head(pfw[altname != "", c("altname", "survey_acronym")])
```

### `wbint_link` {-}

[Minh, could you please explain here?]{style="color:red"}

### `wbext_link` {-}

[Minh, could you please explain here?]{style="color:red"}

### `alt_link` {-}

[Minh, could you please explain here?]{style="color:red"}

### `surv_title` {-}

blah blah

### `surv_producer`  {-}

blah blah

### `survey_coverage`  {-}

This variable represent the househol survey coverage. This is different from the
disaggregation data level.

```{r}
janitor::tabyl(pfw, survey_coverage)
```

### `welfare_type`  {-}

This variable contains the welfare type of the main welfare aggregate variable
in a survey in case it has more than one. The welfare type of alternative
welfare aggregates is found in variable `oth_welfare1_type`.

```{r}
janitor::tabyl(pfw, welfare_type)
```

### `use_imputed`  {-}

Whether the welfare aggregate has been imputed. There are just few countries
with this kind of data.

```{r}
pfw[use_imputed == 1, unique(country_code)]
```

### `use_microdata`  {-}

Whether the welfare aggregate *vector* used in PIP is directly extracted from
microdata without any sort of aggregation. Mos of the countries have this kind
of data. Below you can see those that do not.

```{r}
pfw[use_microdata != 1, unique(country_code)]
```

### `use_bin`  {-}

Whether the welfare aggregate was aggregated to 400 bins from microdata before
being incorporated to the PIP repository. This is the case of houshold surveys
only available in the Luxembourg Data Center
([LIS](https://www.lisdatacenter.org/)). Countries with bin data is considered
micro data for technical purposes.

```{r}
pfw[use_bin == 1, unique(country_code)]
```

### `use_groupdata`  {-}

Whether welfare aggregate comes from grouped data. Information about this type
of data is available in [@dattComputationalToolsPoverty1998;
@krauseCorrigendumEllipticalLorenz2013; @villasenorEllipticalLorenzCurves1989].
The following countries have this kind of data.

```{r}
pfw[use_groupdata == 1, unique(country_code)]
```

### `reporting_year`  {-}
blah blah

### `survey_comparability`  {-}
blah blah

### `comp_note`  {-}
blah blah

### `preferable`  {-}
blah blah

### `show_portal`  {-}
blah blah

### `fieldwork_range`  {-}
blah blah

### `survey_year`  {-}
blah blah

### `newref`  {-}
blah blah

### `ref_year_des`  {-}
blah blah

### `wf_baseprice`  {-}
blah blah

### `wf_baseprice_note`  {-}
blah blah

### `wf_baseprice_des`  {-}
blah blah

### `wf_spatial_des`  {-}
blah blah

### `wf_spatial_var`  {-}
blah blah

### `cpi_replication`  {-}
blah blah

### `cpi_domain`  {-}
blah blah

### `cpi_domain_var`  {-}
blah blah

### `wf_currency_des`  {-}
blah blah

### `ppp_replication`  {-}
blah blah

### `ppp_domain`  {-}
blah blah

### `ppp_domain_var`  {-}
blah blah

### `wf_add_temp_des`  {-}
blah blah

### `wf_add_temp_var`  {-}
blah blah

### `wf_add_spatial_des`  {-}
blah blah

### `wf_add_spatial_var`  {-}
blah blah

### `tosplit`  {-}
blah blah

### `tosplit_var`  {-}
blah blah

### `inpovcal`  {-}
blah blah

### `oth_welfare1_type`  {-}
blah blah

### `oth_welfare1_var`  {-}
blah blah

### `gdp_domain`  {-}
blah blah

### `pce_domain`  {-}
blah blah

### `pop_domain`  {-}
blah blah

### `Note`  {-}
blah blah

### `pfw_id`  {-}
blah blah
